---
title: "COP27"
output:
  html_document:
    toc: true
    toc_float: true
---
<style>
.main-container {
    max-width: 940px;
    margin-left: 0;
    margin-right: auto;
}
</style>

```{r, include=F}
knitr::opts_chunk$set(fig.width=7, fig.height=9, 
                      echo=F, warning=F, message=F)
```

```{r}
library(readr)
library(readxl)
library(dplyr)
library(lubridate)
library(ggplot2)
library(tidyr)
library(flextable)
library(zoo)
library(scales)
library(plotly)

# theme_set(theme_classic() +
#             theme(strip.background=element_rect(linetype=0),
#                   axis.text.x = element_text(angle = 45, hjust=1),
#                   axis.title = element_blank(),
#                   axis.text = element_text(size=12),
#                   axis.line=element_line(colour='grey80'),
#                   axis.ticks = element_line(color="grey80"),
#                   panel.grid.major.y = element_line(color = "grey80"),
#                   plot.title = element_text(size = 17, face = "bold")))
```
```{r}
df = read_csv('../data/netzero.csv')
```

# Introduction
  
How can analysis and visualization of publicly available data be used to further map the ambition of current net-zero CO2eq or GHG non-state actor targets/commitments?  
   
Non-state actors (NSAs) is a term that refers to any group of actors that are not parties to the Paris
Climate Agreement. Cities, subnational governments, and companies–all NSAs–have significant
contributions to climate change mitigation efforts.

Emission reduction targets are a key component of the mitigation ambition of NSAs. This notebook visualizes the progress of NSAs based on the netzero tracker which can be downloaded from https://zerotracker.net.

```{r}
# DATA CLEANING
# Baseline year is incorrect for Releigh, USA
df$end_target_baseline_year[df$end_target_baseline_year==1007 & !is.na(df$end_target_baseline_year)] <- 2007

#End target status
df <- df %>%
  left_join(data.frame(end_target_status=c("Achieved (externally validated)",
                                                          "Achieved (self-declared)",
                                                          "Declaration / pledge",
                                                          "In corporate strategy",
                                                          "In law",
                                                          "In policy document",
                                                          "Proposed / in discussion"),
                                      end_target_status_stage=c(6,5,2,4,4,3,1)))

#Georgia 
if (df$end_target_baseline_year[df$id_code=='COU-0141']>df$end_target_year[df$id_code=='COU-0141']){
  q <- df$end_target_baseline_year[df$id_code=='COU-0141']
  df$end_target_baseline_year[df$id_code=='COU-0141'] <- df$end_target_year[df$id_code=='COU-0141']
  df$end_target_year[df$id_code=='COU-0141'] <- q
}
  
#Poland assume status date is 2021 (assuming data updated till last year)
df$status_date[df$id_code =='COU-0170'] <- 2021

#Angola status year > 2021
df$status_date[df$id_code=='COU-0002'] <- 2021
```

# Emissions reduction targets

## Cities

### Baseline Year vs Target Year
Red - baseline year
Green - status date
Black triangle - target year

```{r}
df %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='City') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(end_target_n_years=end_target_year) %>%
  ggplot(aes(y=name)) +
  geom_point(aes(x=end_target_year), color=1, shape = 2) +
  geom_point(aes(x=end_target_baseline_year), color=2) +
  geom_point(aes(x=status_date), color=3) +
  geom_vline(aes(xintercept=2022), linetype='dashed', color='grey3') -> p
ggplotly(p)
```

```{r}
set.seed(0)
df %>%
  filter(end_target_baseline_year>1800) %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='City') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(end_target_n_years=end_target_year-end_target_baseline_year) %>%
  ggplot(aes(y=name)) +
  geom_point(aes(x=end_target_year-end_target_baseline_year), color=1, shape = 2) +
  geom_point(aes(x=end_target_baseline_year-end_target_baseline_year), color=2) +
  geom_point(aes(x=status_date-end_target_baseline_year), color=3) -> p
ggplotly(p)
```
### % Time elapsed to status date

```{r}
set.seed(0)
percentage_time_elapsed_based_on_status_date <- df %>%
  filter(end_target_baseline_year>1800) %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='City') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(end_target_n_years=end_target_year-end_target_baseline_year) %>%
  mutate(`% time elapsed to target`=(status_date-end_target_baseline_year)/
           (end_target_year-end_target_baseline_year))
  # mutate(name=reorder(name,`% time elapsed to target`)) %>%
percentage_time_elapsed_based_on_status_date %>%
  ggplot(aes(x=name)) +
  geom_col(aes(y=`% time elapsed to target`),
             fill=3) +
  geom_text(aes(y=`% time elapsed to target` + 0.08,
            label=paste(round(`% time elapsed to target`*100,1),'%')), size=2.5) +
  coord_flip()-> p
ggplotly(p)
```

### Status delay (status_date to current year of 2022)

```{r, fig.height=8}
df %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='City') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(status_delay = 2022 - status_date) %>%
  ggplot(aes(x=name)) +
  geom_col(aes(y=status_delay), fill=3) +
  geom_text(aes(y=status_delay + 1,
            label=paste(status_delay,'years')), size=2.5) +
  coord_flip() -> p
ggplotly(p)
```

## Regions

### Baseline Year vs Target Year
Red - baseline year
Green - status date
Black triangle - target year

```{r}
df %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='Region') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(end_target_n_years=end_target_year) %>%
  ggplot(aes(y=name)) +
  geom_point(aes(x=end_target_year), color=1, shape = 2) +
  geom_point(aes(x=end_target_baseline_year), color=2) +
  geom_point(aes(x=status_date), color=3) +
  geom_vline(aes(xintercept=2022), linetype='dashed', color='grey3') -> p
ggplotly(p)
```

```{r}
set.seed(0)
df %>%
  filter(end_target_baseline_year>1800) %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='Region') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(end_target_n_years=end_target_year-end_target_baseline_year) %>%
  ggplot(aes(y=name)) +
  geom_point(aes(x=end_target_year-end_target_baseline_year), color=1, shape = 2) +
  geom_point(aes(x=end_target_baseline_year-end_target_baseline_year), color=2) +
  geom_point(aes(x=status_date-end_target_baseline_year), color=3) -> p
ggplotly(p)
```

### % Time elapsed to status date

```{r}
set.seed(0)
percentage_time_elapsed_based_on_status_date <- df %>%
  filter(end_target_baseline_year>1800) %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='Region') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(end_target_n_years=end_target_year-end_target_baseline_year) %>%
  mutate(`% time elapsed to target`=(status_date-end_target_baseline_year)/
           (end_target_year-end_target_baseline_year))
  # mutate(name=reorder(name,`% time elapsed to target`)) %>%
percentage_time_elapsed_based_on_status_date %>%
  ggplot(aes(x=name)) +
  geom_col(aes(y=`% time elapsed to target`),
             fill=3) +
  geom_text(aes(y=`% time elapsed to target` + 0.08,
            label=paste(round(`% time elapsed to target`*100,1),'%')), size=2.5) +
  coord_flip()-> p
ggplotly(p)
```

### Status delay (status_date to current year of 2022)

```{r, fig.height=8}
df %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='Region') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(status_delay = 2022 - status_date) %>%
  ggplot(aes(x=name)) +
  geom_col(aes(y=status_delay), fill=3) +
  geom_text(aes(y=status_delay + 1,
            label=paste(status_delay,'years')), size=2.5) +
  coord_flip() -> p
ggplotly(p)
```

## Companies

### Baseline Year vs Target Year
Red - baseline year  
Green - status date  
Black triangle - target year  

```{r}
df %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='Company') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(end_target_n_years=end_target_year) %>%
  ggplot(aes(y=name)) +
  geom_point(aes(x=end_target_year), color=1, shape = 2) +
  geom_point(aes(x=end_target_baseline_year), color=2) +
  geom_point(aes(x=status_date), color=3) +
  geom_vline(aes(xintercept=2022), linetype='dashed', color='grey3') -> p
ggplotly(p)
```

```{r}
set.seed(0)
df %>%
  filter(end_target_baseline_year>1800) %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='Company') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(end_target_n_years=end_target_year-end_target_baseline_year) %>%
  ggplot(aes(y=name)) +
  geom_point(aes(x=end_target_year-end_target_baseline_year), color=1, shape = 2) +
  geom_point(aes(x=end_target_baseline_year-end_target_baseline_year), color=2) +
  geom_point(aes(x=status_date-end_target_baseline_year), color=3) -> p
ggplotly(p)
```
### % Time elapsed to status date

```{r}
set.seed(0)
percentage_time_elapsed_based_on_status_date <- df %>%
  filter(end_target_baseline_year>1800) %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='Company') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(end_target_n_years=end_target_year-end_target_baseline_year) %>%
  mutate(`% time elapsed to target`=(status_date-end_target_baseline_year)/
           (end_target_year-end_target_baseline_year))
  # mutate(name=reorder(name,`% time elapsed to target`)) %>%
percentage_time_elapsed_based_on_status_date %>%
  ggplot(aes(x=name)) +
  geom_col(aes(y=`% time elapsed to target`),
             fill=3) +
  geom_text(aes(y=`% time elapsed to target` + 0.08,
            label=paste(round(`% time elapsed to target`*100,1),'%')), size=2.5) +
  coord_flip()-> p
ggplotly(p)
```
### Status delay (status_date to current year of 2022)

```{r, fig.height=8}
df %>%
  filter(end_target=='Emissions reduction target') %>%
  filter(actor_type=='Company') %>%
  # filter(name %in% sample(unique(name),10)) %>%
  mutate(status_delay = 2022 - status_date) %>%
  ggplot(aes(x=name)) +
  geom_col(aes(y=status_delay), fill=3) +
  geom_text(aes(y=status_delay + 1,
            label=paste(status_delay,'years')), size=2.5) +
  coord_flip() -> p
ggplotly(p)
```